
\section{Further studies}

Even though the story we have told in this thesis seems kind of closed, there are many open holes where one can generalize and take inspiration into other areas. I thought I would mention a few things that would have been the next steps for me if there had been an extra year or so to write this thesis. 

There is a natural generalization of DG-algebras to DG-categories, and $\A$-algebras to $\A$-categories. One is also able to define some sort of minimality and formality in these settings, and it would have been interesting to look at them further. There is a relevant category called the Fukaya category, which is a $\A$-category. These are much studied in theoretical physics and homological mirror symmetry, and by [some paper] some subcategories of these categories are formal. 



\subsection*{Massey products on $\A$-algebras}

We mentioned that there is a generalization of Kadeishvili's theorem to $\A$-algebras, meaning that we have for any $\A$-algebra, an $\A$-structure on $H(A)$ and an $\A$-quasi-isomorphism $f\colon H(A)\longrightarrow A$. Stasheff defined Massey products for $\A$-algebras instead of just DG-algebras in \cite{h-spaces}. These were later used in \cite{infty-massey} to prove that the higher products on $H(A)$ do in fact form representatives of the Massey products on an $\A$-algebra $A$. This proof requires that the higher products in $H(A)$ of subsets of the cohomology classes are zero, i.e. that if we want 
\begin{equation*}
    \overline{m_n}(x_1, \ldots, x_n)\in \langle x_1, \ldots, x_n\rangle
\end{equation*}
then we must have $\overline{m_r}(x_{i}, \ldots, x{i+r}) = 0$ for all suitable $i$ and $r$. 

This is also what happens in the DG situation as proved in \cite{detection}, but it looks like the DG situation has some additional room to play with. It's definitely true that the higher products produce representatives of the Massey products when the lower products vanish, as used repeatedly through the thesis, but this can be generalized to certain sets of elements forming defining systems. It would be interesting to check weather this extra room also is possible for the Massey products in $\A$-algebras. 